---
title: "Varying the number of observations"
params: 
  solver: "monolithic"
  n_nodes: 900
  n_replicas: 10
  n_cores: 1
---

## Introduction

In this experiment we're gonna test the behaviour of various types of fpca while varying the number of spatial locations at which the observations are measured

Summing up, we got:
- fixed number of nodes in the mesh, $N$
- varying number of observations, $M$
- fixed number of locations $n = N$

The fpca we're gonna test are a combination of different RegularizedSVD solvers (monolithic and sequential) and different SVD solvers (exact or randomized)
```{r}
svd_versions <- c("exact", "randomized")
```

Recall that:
- The monolithic approach involves a Cholesky decomposiyion of a matrix of dimensions $N^2 \times N^2$ and a single SVD of a matrix of dimensions $M \times N^2$

- The sequential approach involves the solution of a single SVD of a matrix of dimensions $M \times n^2$ and the sequential update of the components

```{r}
library(fdaPDE)
library(femR)

mesh_data_path <- "../data/mesh/unit_square"
obs_data_path <- "../data/models/fpca/2D_test"
results_path <- "results"

plot_path <- "plots-n_unitsVAR"

n_replicas <- params$n_replicas

file.remove(list.files(path = mesh_data_path, full.names = TRUE))
file.remove(list.files(path = obs_data_path, full.names = TRUE))
file.remove(list.files(path = results_path, full.names = TRUE))
```

Set the hyperparameters for the experiment by changing the file `params.json` in the folder 
```{r}
library(jsonlite)

json_data <- fromJSON("params.json")

json_data$RunParams$RSVDType <- unbox("sequential")
json_data$RunParams$SVDType <- unbox("randomized")

json_data$RunParams$n_replicas <- unbox(params$n_replicas)
json_data$RunParams$lambda <- unbox(params$lambda)

write_json(json_data, "params.json", pretty = TRUE)
```

- set the parameters to generate the data
```{r}
n_units <- params$n_statistical_units
n_replicas <- params$n_replicas

seed = 467897965 ## for reproducibility purposes
```

## Data
```{r}
source("../r-utils/mesh_utils.R")
generate.mesh_and_eigenfunctions <- function(n_nodes, mesh_path,
                                    n_locs, locs_path){
  ##(1) Mesh generation
  mesh.2D <- unit_square.mesh(n_nodes, mesh_path)
  #'define a dummy pde to get the stiffness and the mass matrices of the 
  #'differential operator
  mesh <- Mesh(
    list(
      nodes = mesh.2D$nodes,
      elements = mesh.2D$triangles,
      boundary = mesh.2D$nodesmarkers
    )
  )
  Vh <- FunctionSpace(mesh, fe_order=1)
  u <- Function(Vh)
  
  L <- anisothropic_diffusion_op(
    pi/3, #angle
    1.5 #intensity
    )
  f <- function(points) { return(0*points[,1])} #forcing term set to 0 
  pde <- Pde(L(u), f)
  
  ##(2) Eigenfunctions
  #evd
  evd <- eigen(Pde(L(u),f)$stiff(), symmetric = T, only.values = F)
  #at mesh nodes
  f1_fem <- evd$vectors[,20]
  f2_fem <- evd$vectors[,29]
  f3_fem <- evd$vectors[,4]

  ##(3) Evaluate at locations
  locations.obs.2D <- unit_square.locs(n_locs,locs_path) #locations at nodes
  Psi <- Psi.matrix(mesh.2D, locations.obs.2D)
  #L2 norm
  L2norm <- L2norm.factory(mesh.2D, L) #function to compute the norm
  #final components
  f1 <- as.matrix(Psi %*% f1_fem/L2norm(f1_fem))
  f2 <- as.matrix(Psi %*% f2_fem/L2norm(f2_fem))
  f3 <- as.matrix(Psi %*% f3_fem/L2norm(f3_fem))
  
  return(cbind(f1,f2,f3))
}
```

Actual generation (mesh & PCs)
```{r}
true_PCs <- generate.mesh_and_eigenfunctions(
  params$n_nodes, mesh_data_path,
  params$n_nodes, obs_data_path #locsEQnod
)
```

## Run
```{r}
test_nobs <- c(50,100,200,300,400,600,800,1600)

seed = 467897965 ## for reproducibility purposes
```

To store the results
```{r}
library(dplyr)

keys <- c("rsvd","svd","replica","n_units")

error_metrics <- c("errPC1","errPC2","errPC3",
                   "errScore1","errScore2","errScore3",
                   "signal_reconstruction_error",
                   "space_reconstruction_error")

time_metrics <- c("exe_time","svd_time","other_time")

test_results <- 
   data.frame(matrix(nrow=0,ncol=length(c(keys,error_metrics,time_metrics))))
names(test_results) <- c(keys, error_metrics,time_metrics)

test_results[] <- lapply(test_results, function(x) numeric())

test_results$rsvd <- character()
test_results$svd <- character()
```

```{r}
source("../r-utils/error_metrics.R")


for(k in seq_along(test_nobs)){
  ### number of statistical units
  N <- test_nobs[k]
  cat("\nNumber of observations: ",N,"\n")

  for(ind_rep in 1:n_replicas) {
    data <- 
      data_generator(N,
                     true_PCs = true_PCs,
                     scores_sd = c(0.4,0.3,0.2),
                     error_sd = 0.1,
                     obs_path = obs_data_path,
                     seed = seed + 4*ind_rep)
    
    #repeat the experiment for each svd_version
    for(rsvd_version in c("sequential","monolithic")){
      cat("RSVD version: ", rsvd_version, "\n")
      json_data$RunParams$RSVDType <- unbox(rsvd_version)
      for(svd_version in svd_versions){
        cat("Svd version: ", svd_version, "\n")
        
        #modify the params.json
        json_data$RunParams$SVDType <- unbox(svd_version)
        write_json(json_data, "params.json", pretty = TRUE)
        
        #run the program
        system("/Users/marcogalliani/Desktop/fdaPDE-cpp/cmake-build-default/experiments/fpca/fpca_test
    ")
        
        #results
        est.PCs <- as.matrix(readMM(file=paste(results_path,"estimated_PCs.mtx",sep="/")))
        est.Scores <- as.matrix(readMM(file=paste(results_path,"estimated_Scores.mtx",sep="/")))
        exe_time <- read.csv(file=paste(results_path,"execution_times.csv",sep="/"),header=F)[[1]][1]
        time_svd <- read.csv(file=paste(results_path,"svd_time.csv",sep="/"),header=F)[[1]][1]
        time_other <- exe_time - time_svd
        
        #compute errors & times
        errors <- compute.errors.FPCA(data, 
                                      est.PCs = est.PCs,
                                      est.Scores = est.Scores
                                      )
        
        test_results <- test_results %>%
          add_row(rsvd=rsvd_version, svd=svd_version, replica=ind_rep, n_units=N)
        new_row <- 
          test_results$rsvd==rsvd_version & 
          test_results$svd==svd_version & 
          test_results$replica==ind_rep &
          test_results$n_units==N
        test_results[new_row, error_metrics] <- unlist(errors)
        test_results[new_row, time_metrics] <- c(exe_time,time_svd,time_other)
      }#svd version
    }#rsvd version
  }#replicas
}#n_obs
```


## Saving the results
```{r}
write.csv(format(test_results, digits = 16),
          paste("past-runs",
                paste("overall-fpca-test",
                      "-Mesh",params$n_nodes,
                      "-Obs","VAR",
                      "-LocsEQnod",
                      ".csv",sep=""),
          sep="/")
)
```

