---
title: "Varying the number of nodes"
params: 
  n_statistical_units: 400 #[200,400,800,"VAR"]
  n_replicas: 10
  n_cores: 1
---

## Introduction

In this experiment we're gonna test the behaviour of various types of fpca while varying the number of spatial locations at which the observations are measured

Summing up, we got:
- varying number of nodes in the mesh, $N$
- fixed number of observations, $M$
- varying number of locations, s.t. $n = N$

The fpca we're gonna test are a combination of different RegularizedSVD solvers (monolithic and sequential) and different SVD solvers (exact or randomized)
```{r}
rsvd_versions <- c("simple_monolithic", "spchol_monolithic")
```


Recall that:
- The monolithic approach involves a Cholesky decomposiyion of a matrix of dimensions $N^2 \times N^2$ and a single SVD of a matrix of dimensions $M \times N^2$

- The sequential approach involves the solution of a single SVD of a matrix of dimensions $M \times n^2$ and the sequential update of the components

```{r}
library(fdaPDE)
library(femR)

mesh_data_path <- "../data/mesh/unit_square"
obs_data_path <- "../data/models/fpca/2D_test"
results_path <- "results"
plot_path <- "plots/varying-mesh"

n_replicas <- params$n_replicas

file.remove(list.files(path = mesh_data_path, full.names = TRUE))
file.remove(list.files(path = obs_data_path, full.names = TRUE))
file.remove(list.files(path = results_path, full.names = TRUE))
```

Set the hyperparameters for the experiment by changing the file `params.json` in the folder 
```{r}
library(jsonlite)

json_data <- fromJSON("params.json")

json_data$RunParams$RSVDType <- unbox("spchol_monolithic")
json_data$RunParams$n_cores <- unbox(params$n_cores)

json_data$RunParams$SVDType <- unbox("exact")
json_data$RunParams$lambda <- unbox(1e-2)

write_json(json_data, "params.json", pretty = TRUE)
```


## Mesh
```{r}
test_nnodes <- c(100,400,900,1600,3600,4900)

seed = 467897965 ## for reproducibility purposes
```


```{r}
test_results <- 
   data.frame(rsvd = rep(rep(rsvd_versions, each=n_replicas),times=length(test_nnodes)),
              n_nodes = as.factor(rep(test_nnodes,each=n_replicas*length(rsvd_versions))),
              replica = as.factor(rep(1:n_replicas, times=length(test_nnodes))))

error_metrics <- c("errPC1","errPC2","errPC3",
                   "errScore1","errScore2","errScore3",
                   "signal_reconstruction_error",
                   "space_reconstruction_error")
test_results[,error_metrics] <- NA

time_metrics <- c("exe_time","svd_time","other_time")
test_results[,time_metrics] <- NA
```


```{r}
library(pracma)
source("../include/mesh_utils.R")
source("../include/error_metrics.R")
```

```{r}
for(k in seq_along(test_nnodes)){
  
  ##(1) Mesh generation
  N <- test_nnodes[k] ## number of nodes in spatial domain
  cat("\nNumber of Nodes: ", N,"\n")
  
  mesh.2D <- unit_square.mesh(N,mesh_data_path)
  
  ##(2) Eigenfunctions
  ## use n = N for the setting locations = mesh nodes
  ## (in this case, \Psi is the identity matrix)
  ## use n != N for the general case. Set n > N.
  n <- N
  locations.obs.2D <- unit_square.locs(n,obs_data_path)
  n_locations <- nrow(locations.obs.2D)
  
  L2norm <- L2norm.factory(mesh.2D) #function to compute the norm
  
  ## (2.1) at mesh nodes
  ## evaluate eigenfunctions on mesh nodes (this is equivalent to get the FEM basis
  ## expansion of the eigenfunctions ). change values of a and b to select eigenfunction
  f1_fem <- square_eigenfunction(a = 1, b = 1, mesh.2D$nodes)
  f2_fem <- square_eigenfunction(a = 2, b = 3, mesh.2D$nodes)
  f3_fem <- square_eigenfunction(a = 4, b = 4, mesh.2D$nodes)

  ## (2.2) at locations (normalized)
  Psi <- Psi.matrix(mesh.2D, locations.obs.2D)
  

  f1 <- as.matrix(Psi %*% f1_fem/L2norm(f1_fem))
  f2 <- as.matrix(Psi %*% f2_fem/L2norm(f2_fem))
  f3 <- as.matrix(Psi %*% f3_fem/L2norm(f3_fem))
  
  true_PCs <- cbind(f1,f2,f3)
  
  ##(3) Data generation
  M <- params$n_statistical_units ##fixed
  
  for(i in 1:n_replicas) {
    cat("Replica: ", i, "\n")

    data <- data_generator(M,
                           true_PCs = true_PCs,
                           scores_sd = c(0.4,0.3,0.2),
                           error_sd = 0.1,
                           obs_path = obs_data_path,
                           seed = seed + 4*i)
    
    #repeat the experiment for each svd_version
    for(rsvd_v in rsvd_versions){
      cat("rsvd version: ", rsvd_v, "\n")
      
      #modify the params.json
      json_data$RunParams$RSVDType <- unbox(rsvd_v)
      write_json(json_data, "params.json", pretty = TRUE)
      
      #run the program
      system("/Users/marcogalliani/Desktop/fdaPDE-cpp/cmake-build-default/experiments/sparsity-monolithic/sparse_mono_fpca --gtest_filter=* --gtest_color=no")
      
      #load results
      getwd()
      est.PCs <- 
        as.matrix(readMM(file=paste(results_path,"estimated_PCs.mtx",sep="/")))
      est.Scores <- 
        as.matrix(readMM(file=paste(results_path,"estimated_Scores.mtx",sep="/")))
      
      #compute errors & times
      errors <- list()
      errors$errPC <- vector()
      errors$errScore <- vector()
      
      for(ind_pc in 1:3){
        errors$RMSE.PCs[ind_pc] <- min(RMSE.vector(true_PCs[,ind_pc],est.PCs[,ind_pc]),
                                  RMSE.vector(true_PCs[,ind_pc],-est.PCs[,ind_pc]))
        errors$Scores[ind_pc] <- min(RMSE.vector(data$scores[,ind_pc],est.Scores[,ind_pc]),
                                RMSE.vector(data$scores[,ind_pc],-est.Scores[,ind_pc]))
      }
      errors$signal_reconstruction <- RMSE.matrix(data$data_matrix,est.Scores%*%t(est.PCs))
      errors$space_reconstruction <- subspace(true_PCs,est.PCs)
  
      times <- list()
      times$exe_time <- read.csv(file=paste(results_path,"execution_times.csv",sep="/"),
                                 header=F)[[1]][1]
      times$svd_time <- read.csv(file=paste(results_path,"time_svd.csv",sep="/"),
                                 header=F)[[1]][1]
      times$other_time <- read.csv(file=paste(results_path,"time_other.csv",sep="/"),
                                   header=F)[[1]][1]
      
      # write in test result
      current_run <- 
        test_results$rsvd ==rsvd_v & 
        test_results$n_nodes==N & 
        test_results$replica==i
      
      test_results[current_run, c(error_metrics,time_metrics)] <-
        c(unlist(errors),unlist(times))
    }#svd version
  }#replica
}#n_nodes
```

## Saving the results
```{r}
write.csv(format(test_results, digits = 16),
          paste("past-runs",
                paste(json_data$RunParams$RSVDType,"-MeshVAR","-Obs",M,"-LocsEQnod",
                         ".csv",sep=""),
          sep="/")
)
```