---
title: "Base run"
params: 
  solver: "monolithic"
  n_nodes: 20 #10, 20, 30, 40, 60
  n_obs: 200
  n_locs: 30
  n_cores: 1
---

## Introduction

In this experiment we're gonna test the behaviour of various types of fpca while varying the number of spatial locations at which the observations are measured

The fpca we're gonna test are a combination of different RegularizedSVD solvers (monolithic and sequential) and different SVD solvers (exact or randomized)
```{r}
svd_versions <- c("exact", "rsi", "rbki")
```

Recall that:
- The monolithic approach involves a Cholesky decomposiyion of a matrix of dimensions $N^2 \times N^2$ and a single SVD of a matrix of dimensions $M \times N^2$

- The sequential approach involves the solution of a single SVD of a matrix of dimensions $M \times n^2$ and the sequential update of the components

```{r}
library(fdaPDE)
library(femR)

mesh_data_path <- "../data/mesh/unit_square"
obs_data_path <- "../data/models/fpca/2D_test"
results_path <- "results"
plot_path <- "plots/varying-obs"

file.remove(list.files(path = mesh_data_path, full.names = TRUE))
file.remove(list.files(path = obs_data_path, full.names = TRUE))
file.remove(list.files(path = results_path, full.names = TRUE))
```

Set the hyperparameters for the experiment by changing the file `params.json` in the folder 
```{r}
library(jsonlite)

json_data <- fromJSON("params.json")

json_data$RunParams$RSVDType <- unbox(params$solver)
json_data$RunParams$n_replicas <- unbox(1)
json_data$RunParams$n_cores <- unbox(params$n_cores)

json_data$SVDParams$SVDType <- unbox("exact")
json_data$SVDParams$tolerance <- unbox(1e-2)

json_data$RSI$oversampling <- unbox(3)
json_data$RBKI$block_size <- unbox(10)
```


## Mesh
```{r}
## define spatial domain
N <- params$n_nodes ## number of nodes in spatial domain
x.2D <- seq(0, 1, length.out = N)
y.2D <- x.2D
locations.2D <- expand.grid(x.2D, y.2D)
mesh.2D <- fdaPDE::create.mesh.2D(locations.2D)

## export mesh
write.csv(format(mesh.2D$nodes, digits = 16), paste(mesh_data_path,"points.csv", sep = "/"))
write.csv(format(mesh.2D$triangles, digits = 16), paste(mesh_data_path,"elements.csv", sep = "/"))
write.csv(format(1 * mesh.2D$nodesmarkers, digits = 16), paste(mesh_data_path,"boundary.csv", sep = "/"))
write.csv(format(mesh.2D$neighbors, digits = 16), paste(mesh_data_path,"neigh.csv", sep = "/"))
write.csv(format(mesh.2D$edges, digits = 16), paste(mesh_data_path,"edges.csv", sep = "/"))
```

Extracting the mass matrix to normalize the eigenfunction later on
```{r}
## use femR to extract mass matrix
mesh <- Mesh(
  list(
    nodes = mesh.2D$nodes,
    elements = mesh.2D$triangles,
    boundary = mesh.2D$nodesmarkers
  )
)
## define a dummy PDE, just to get the mass matrix
Vh <- FunctionSpace(mesh, fe_order=1)
u <- Function(Vh)
Lu <- -laplace(u) ## poisson problem
f <- function(points) { return(0*points[,1])}
pde <- Pde(Lu, f)

R0 <- pde$mass() ## mass matrix
```

## Eigenfunctions
Definition of the eigenfunctions
```{r}
## define eigenfunctions of laplace operator over square with neumann boundary conditions
square_eigenfunction <- function(a, b, locs) {
    return(cos(a * pi * locs[,1]) * cos(b * pi * locs[, 2]))
}
```

Evaluation at the nodes and normalization
```{r}
L2norm <- function(f) { return(sqrt(as.numeric(t(f) %*% R0 %*% f))) }

## evaluate eigenfunctions on mesh nodes (this is equivalent to get the FEM basis
## expansion of the eigenfunctions ). change values of a and b to select eigenfunction
f1_fem <- square_eigenfunction(a = 1, b = 1, mesh.2D$nodes)
f2_fem <- square_eigenfunction(a = 2, b = 3, mesh.2D$nodes)
f3_fem <- square_eigenfunction(a = 4, b = 4, mesh.2D$nodes)

## normalize
f1_norm <- L2norm(f1_fem)
f2_norm <- L2norm(f2_fem)
f3_norm <- L2norm(f3_fem)

f1_normalized <- f1_fem / f1_norm
f2_normalized <- f2_fem / f2_norm
f3_normalized <- f3_fem / f3_norm
```


## Locations
```{r}
## use n = N for the setting locations = mesh nodes
## (in this case, \Psi is the identity matrix)
## use n != N for the general case. Set n > N.

n <- params$n_locs

## define observations' locations
x.obs.2D <- seq(0, 1, length.out = n)
y.obs.2D <- x.obs.2D
locations.obs.2D <- expand.grid(x.obs.2D, y.obs.2D)
n_locations <- nrow(locations.obs.2D)
  
## export locations
write.csv(format(locations.obs.2D, digits = 16), paste(obs_data_path,"locs.csv", sep = "/"))
```

Evaluation of the eigenfunctions at the locations
```{r}
## obervations are evaluation of L^2 normalized eigenfunctions at data locations
Psi <- Vh$basis()$eval(as.matrix(locations.obs.2D))

f1 <- as.matrix(Psi %*% f1_normalized)
f2 <- as.matrix(Psi %*% f2_normalized)
f3 <- as.matrix(Psi %*% f3_normalized)
```


## Datamatrix
```{r}
seed = 467897965 ## for reproducibility purposes
```



```{r}
library(pracma)
library(Matrix)



M <- params$n_obs
cat("\nNumber of observations: ",M,"\n")

## set score standard deviation
sd_score1 <- 0.4
sd_score2 <- 0.3
sd_score3 <- 0.2

sd_error  <- 0.1

data_range <- max(c(f1, f2, f3)) - min(c(f1, f2, f3))

set.seed(seed)
## sample scores from normal distribution
score1 <- rnorm(n = M, sd = sd_score1 * data_range)
score2 <- rnorm(n = M, sd = sd_score2 * data_range)
score3 <- rnorm(n = M, sd = sd_score3 * data_range)
## generate true observations
datamatrix_pointwise_exact <-
    matrix(score1) %*% t(matrix(f1)) +
    matrix(score2) %*% t(matrix(f2)) +
    matrix(score3) %*% t(matrix(f3))

## add error to simulate real data
error <- rnorm(n = M * n_locations, sd = sd_error * data_range)
datamatrix_pointwise <- datamatrix_pointwise_exact + error

## datamatrix.pointwise are the observations
## (pointwise) center data
data_bar <- colMeans(datamatrix_pointwise)
data_bar <- matrix(rep(data_bar, nrow(datamatrix_pointwise)),
                   nrow = nrow(datamatrix_pointwise),
                   byrow = TRUE)

datamatrix_pointwise_centred <- datamatrix_pointwise - data_bar

## export data .csv format
write.csv(format(datamatrix_pointwise_centred, digits = 16),
          paste(obs_data_path,"datamatrix_centred.csv",sep="/")
)
```

## Testing with different SVD solvers
```{r}
#repeat the experiment for each svd_version
estimated_PCs <- list()

for(svd_version in svd_versions){
  cat("Svd version: ", svd_version, "\n")
  
  #modify the params.json
  json_data$SVDParams$SVDType <- unbox(svd_version)
  write_json(json_data, "params.json", pretty = TRUE)
  
  #run the program
  system("/Users/marcogalliani/Desktop/fdaPDE-cpp/cmake-build-default/experiments/fpca/fpca_test --gtest_filter=* --gtest_color=no")
  
  #results
  estimated_PCs[[svd_version]] <- 
    as.matrix(readMM(file=paste(results_path,"estimated_PCs.mtx",sep="/")))
}#svd version
```


## Visualization
```{r}
library(ggplot2)
library(viridis)

plot_eigenfunction <- function(data, locs, title) {
    ## define data.frame for plotting
    df <- data.frame(
        x = locs[,1],
        y = locs[,2],
        f = data
    )
    ## colors
    n_breaks <- 50
    mybreaks <- c(-Inf, seq(min(data), max(data), length.out = n_breaks), Inf)
    mycolors<- function(x) {
        colors<-colorRampPalette(viridis(11))( x + 1 )
        colors[1:x]
    }
    ## plot
    p <- ggplot() +
        geom_contour_filled(data = df, aes(x, y, z = f),
                            breaks = mybreaks) +
        scale_fill_manual(
            aesthetics = "fill",
            values = mycolors(n_breaks + 2), name = "Value", drop = FALSE
        ) +
        coord_equal() + 
        theme_void() +
        theme(legend.position = "none",plot.title = element_text(hjust = 0.5)) + 
        ggtitle(title)
  
    p
}
```

```{r}
library(Matrix)
library(gridExtra)

plot_PCs <- list()


for(i in 1:3){
  plot_PCs[[i]] <- list()
  for(svd in svd_versions){
    plot_PCs[[i]][[svd]] <- plot_eigenfunction(-estimated_PCs[[svd]][,i], 
                                     locations.obs.2D, svd)
  }
}

PCs <- list(f1,f2,f3)

grid_PCs <- list()
for(i in 1:3){
  grid_PCs[[i]] <- arrangeGrob(
    plot_eigenfunction(PCs[[i]], locations.obs.2D,"original"), 
    arrangeGrob(
        plot_PCs[[i]][["exact"]], plot_PCs[[i]][["rsi"]],
        plot_PCs[[i]][["rbki"]], plot_PCs[[i]][["rbki_ext"]]
       ),
     ncol=2, nrow=1, top = paste("PC",i,sep=""))
}
```

```{r}
library(gridExtra)

for(i in 1:3){
  ggsave(paste("plots/example-PCs/",
               "PC",i,
               "-",params$solver,
               "-Mesh",params$n_nodes,
               "-Obs",params$n_obs,
               "-Locs",params$n_locs,
             ".pdf", sep=""), 
       grid_PCs[[i]], width = 10, height = 5)
}
```





